---
title: "Final project"
author: "Mandi Ward & Marielena McWhirter"
date: "5/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      eval = TRUE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.width = 5,
                      fig.height = 4.25)

library(tidyverse)
library(afex)
library(rio)
library(here)
library(knitr)
library(kableExtra)
library(misty)
library(janitor)
library(rstatix)
library(dplyr)
library(psych)
library(MOTE)

theme_set(theme_minimal())
```

2 x 3 factorial ANOVA:
IV: Dating, married, or single
IV: living in a house or living in a dorm

DV: BDI scores

table() function to look at grouping 
with(df, table(var1, var2))
table(df$var1, df$var2)


BELOW IS WHERE I STARTED OVER TRYING TO CREATE NEW VARIABLE


THIS CODE WORKS
# Load in the data
```{r}
act_t1 <- import(here("data", "ACT Suicide Prevention T1.sav"), setclass = "tbl_df") %>% 
  characterize() %>% 
  clean_names()

act_t1_12520 <- import(here("data", "ACT Suicide Prevention T1 1-25-20.sav"), setclass = "tbl_df") %>% 
  characterize() %>% 
  clean_names()
```

# Joining datasets
```{r}
act_all <- full_join(act_t1, act_t1_12520)
```

# Creating df with our variables of interest
```{r}

df <- act_all %>% 
  select(id, dem1_17a, dem1_02a, bdi1)

df_1 <- df %>% 
  drop_na(dem1_17a, dem1_02a, bdi1)

df_2 <- df_1[-c(220, 307, 651),]

df_3 <- df_2[-c(220),]

df_4 <- df_3 %>% 
  rename(livingstat = dem1_02a,
         datingstat = dem1_17a,
         BDI = bdi1)
```

# Describing Data

```{r}
describe(df)
```





# Frequency table 
```{r}
df_4 %>% 
  group_by(livingstat, datingstat) %>% 
  summarise(n = n()) %>% #new column "n" = row count of each factor grouping
  spread(livingstat, n) # rows "spread" across the top
```

# Checking the assumptions with descriptives
Descriptives:
  The average BDI score is 7.68 (6.99), and IS NOT normally distributed (skew=1.77, kurtosis=4.44).
```{r}
describe(df_4)

describeBy(df_4, group = "livingstat")
```

# Visually inspecting the data 
  Step 1: Extract means to examine group differences
```{r}
means <- df_4 %>% #creating new object (tibble) with the means
  group_by(livingstat, datingstat) %>% #grouping our results by our two independent variables, 'country' within levels of 'sen2'
  summarise(mean_bdi = mean(BDI))
means
```

Step 2: Examine the means using a profile plot
```{r}
means %>% 
  ggplot(aes(livingstat, mean_bdi, group = datingstat)) +
  geom_point() +
  geom_line(aes(color = datingstat)) +
  labs(title = "Profile Plot of Means", x = "Living Location", y = "BDI Score", fill = "Relationship Status")
```

 Visualize using boxplots
```{r}
df_4 %>% 
  ggplot(aes(x = datingstat, y = BDI, fill = livingstat)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="bottom") +
  labs(title = "Boxplot of Means", x = "Relationship Status", y = "BDI Score", fill = "Living Location")

```
 
### Assessing for homogeneity of variance with a Levene's Test.
  Levene's Test on the mean of the variables is significant. This indicates that the variables do not meet the assumption of homogeneity in variance. However, experiemental study designs have more flexability with Levene's Test, and homogeneity in variance is not required to be considered for ANOVA analyses.
```{r}
car::leveneTest(BDI ~ livingstat*datingstat, data = df_4, center = "mean")
```

## ANOVA 
```{r}
### THIS DOES NOT WORK
#anova_test(df_3, formula = bdi1 ~ dem1_02a + dem1_17a + dem1_02a:dem1_17a, 
                 #detailed = TRUE, type = 3, effect.size = "pes")


## THIS ONE WORKS
res.aov2 <- aov(BDI ~ livingstat + datingstat + livingstat:datingstat, data = df_4)
summary(res.aov2)
```

## Effect size for interaction
```{r}
 omega.F(dfm = 1, dfe = 4,
      Fvalue = 0.073, n = 963, a = .05)
```

## Effect size for main effect
```{r}
omega.F(dfm = 1, dfe = 404,
      Fvalue = 8.305, n = 963, a = .05)
```

ADD IN QQ PLOT VISUAL IN ADDITION TO SKEW AND KURTOSIS 
in discussion discuss implications and not meeting criteria and just that it may impact the validity of the results.

```{r}
df %>% 
  ggplot(aes(sample = bdi1)) +
    stat_qq()+
    stat_qq_line(color = "violet") +
    theme_classic() +
    labs(title = "QQ Plot of Mean Depression Scores")
```

